Competing Bandits: \\ The Perils of Exploration under Competition

### 2019-02-14 paper/content/non_greedy_choice.tex
add HardMax macro

In this section we consider an extension of our base model where the agents' choice model is no longer deterministic. Recall that in our base model agents deterministically choose the firm with the higher reputation score. We define this choice model as \textit{HardMax}. We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the higher reputation score with probability $1 - \epsilon$. Following \cite{mansour2018competing} we define this choice model as \textit{HardMax with random} (\HMR).

@@

\footnotesize
\begin{table*}[t]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
  \hline
  & \multicolumn{3}{c||}{Heavy-Tail (\HMR)}
  & \multicolumn{3}{c|}{Heavy-Tail (HardMax)}\\
  \hline
  & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG 
 & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG  \\
  \hline
$T = 2000$
 & \makecell{ \textbf{0.43} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.44} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.02 \\Var: 0.1 }
  %%
 &  \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.28} $\pm$ 0.03 \\Var: 0.19 } 
  & \makecell{ \textbf{0.63} $\pm$ 0.03 \\Var: 0.18 }
    \\
\hline
  $T= 5000$ 
   & \makecell{ \textbf{0.66} $\pm$ 0.01 \\Var: 0.056 } 
  & \makecell{ \textbf{0.59} $\pm$ 0.02 \\Var: 0.092 } 
  & \makecell{ \textbf{0.56} $\pm$ 0.02 \\Var: 0.098 } 
  %%
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.62} $\pm$ 0.03 \\Var: 0.19 }
 \\ 
  \hline
  $T = 10000$
  & \makecell{ \textbf{0.76} $\pm$ 0.01 \\Var: 0.026 } 
 & \makecell{ \textbf{0.67} $\pm$ 0.02 \\Var: 0.067 } 
 & \makecell{ \textbf{0.52} $\pm$ 0.02 \\Var: 0.11 }
 %%
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.21 } 
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.03 \\Var: 0.2 }
  %%
  \\
   \hline
\end{tabular}
\normalsize
\caption{Performance Comparison between HardMax and \HMR Choice on the Heavy-Tail MAB instance. Each cell describes the market shares in a game between two algorithms, call them Alg1 vs. Alg2, at a particular value of $t$. Line 1 in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the \HMR choice model and $T_0 = 20$.}
\label{tab:non_greedy_table}
\end{table*}

@@

Table \ref{tab:non_greedy_table} shows the average market shares under the HardMax vs \HMR choice model. As $T$ gets sufficiently large, \TS becomes weakly dominant for the \HMR model whereas, as expected, it does not for the HardMax model. These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$). However, they are drawn from the same prior instances and so qualitatively are the same but the quantitative results are not directly comparable to those from the previous section.} The intuition for why \TS should become the dominant strategy eventually is simple. The existence of the random agents ensures that no algorithm ever gets completely starved of agents. As a result, each algorithm should eventually converge to its asymptotic performance in isolation.

However, it takes a significant amount of randomness and a relatively large time horizon for this convergence to occur. Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance. Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that \HMR leads to lower variance in market shares across simulations compared to HardMax.

>>>>

In this section we consider an extension of our base model where the agents' choice model is no longer deterministic. Recall that in our base model agents deterministically choose the firm with the higher reputation score. We define this choice model as \textit{HardMax} (\HM). We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the higher reputation score with probability $1 - \epsilon$. Following \cite{mansour2018competing} we define this choice model as \textit{HardMax with random} (\HMR).

@@

\footnotesize
\begin{table*}[t]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
  \hline
  & \multicolumn{3}{c||}{Heavy-Tail (\HMR)}
  & \multicolumn{3}{c|}{Heavy-Tail (\HM)}\\
  \hline
  & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG 
 & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG  \\
  \hline
$T = 2000$
 & \makecell{ \textbf{0.43} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.44} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.02 \\Var: 0.1 }
  %%
 &  \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.28} $\pm$ 0.03 \\Var: 0.19 } 
  & \makecell{ \textbf{0.63} $\pm$ 0.03 \\Var: 0.18 }
    \\
\hline
  $T= 5000$ 
   & \makecell{ \textbf{0.66} $\pm$ 0.01 \\Var: 0.056 } 
  & \makecell{ \textbf{0.59} $\pm$ 0.02 \\Var: 0.092 } 
  & \makecell{ \textbf{0.56} $\pm$ 0.02 \\Var: 0.098 } 
  %%
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.62} $\pm$ 0.03 \\Var: 0.19 }
 \\ 
  \hline
  $T = 10000$
  & \makecell{ \textbf{0.76} $\pm$ 0.01 \\Var: 0.026 } 
 & \makecell{ \textbf{0.67} $\pm$ 0.02 \\Var: 0.067 } 
 & \makecell{ \textbf{0.52} $\pm$ 0.02 \\Var: 0.11 }
 %%
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.21 } 
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.03 \\Var: 0.2 }
  %%
  \\
   \hline
\end{tabular}
\normalsize
\caption{Performance Comparison between \HM and \HMR Choice on the Heavy-Tail MAB instance. Each cell describes the market shares in a game between two algorithms, call them Alg1 vs. Alg2, at a particular value of $t$. Line 1 in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the \HMR choice model and $T_0 = 20$.}
\label{tab:non_greedy_table}
\end{table*}

@@

Table \ref{tab:non_greedy_table} shows the average market shares under the \HM vs \HMR choice model. As $T$ gets sufficiently large, \TS becomes weakly dominant for the \HMR model whereas, as expected, it does not for the \HM model. These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$). However, they are drawn from the same prior instances and so qualitatively are the same but the quantitative results are not directly comparable to those from the previous section.} The intuition for why \TS should become the dominant strategy eventually is simple. The existence of the random agents ensures that no algorithm ever gets completely starved of agents. As a result, each algorithm should eventually converge to its asymptotic performance in isolation.

However, it takes a significant amount of randomness and a relatively large time horizon for this convergence to occur. Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance. Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that \HMR leads to lower variance in market shares across simulations compared to \HM.
